---
title: "Statistical Analysis"
output:
  html_document: default
  pdf_document: default
---

# Analysis of results

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(broom)
library(lme4)
library(nnet)
library(xtable)
library(WebPower)
library(ggplot2)
library(gridExtra)
library(ggsignif)
library(dplyr)
library(tidyr)
library(sjPlot)
library(interactions)
library(logistf)

# Clear environment
rm(list=ls())

#### Set the theme of the plots
theme_set(theme_bw()+ 
            theme(plot.title = element_text(colour = "black", face = "bold",hjust = 0.5)) + theme(axis.text = element_text(size = 12, colour = "black", face = "bold"))+
            theme(axis.title = element_text(size = 13, colour = "black", face = "bold")))
```

## Loading data and encode variables

```{r}
# Load the integrated CSV file, and you should change the name of file
data <- read.csv('integrated_result_gpt5nano.csv', stringsAsFactors = FALSE)

data <- data %>%
  rename(
    id = Subject,
    item = Item_ID,
    subexp = Group,
    accuracy_num = Is_Correct
  )

data$subexp <- factor(data$subexp, levels = c("BAS", "DIST", "UPP", "ESQ"))

data$trial_index <- as.numeric(data$trial_index)

# For BAS, we assign a placeholder or keep NA initially
data$condition_factor <- NA
data$condition_factor[data$Prev_Prime_Cat == 'Critical'] <- "critical"
data$condition_factor[data$Prev_Prime_Cat == 'Control'] <- "control"
data$condition_factor[data$Prev_Prime_Cat == 'BAS'] <- "baseline"

# Critical=1, Control=0
data$condition <- NA
data$condition[data$Prev_Prime_Cat == 'Critical'] <- 1
data$condition[data$Prev_Prime_Cat == 'Control'] <- 0
```

## Filter data subset

```{r}
fillers <- filter(data, Type == 'Filler')
data_exp <- filter(data, Type %in% c('Prime', 'Target'))
data_exp$is_prime <- data_exp$Type == 'Prime'
```

## Accuracy in fillers

```{r}
# Overall filler accuracy
mean(fillers$accuracy_num)

# Filler accuracy by sub-experiment
fillers %>%
   group_by(subexp) %>%
  dplyr::summarise(
    n=n(),
    mean=mean(accuracy_num),
    sd=sd(accuracy_num)
  )

# Plot Filler Accuracy
fillers %>%
   group_by(subexp) %>%
  dplyr::summarise(
    n=n(),
    mean=mean(accuracy_num),
    sd=sd(accuracy_num)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  # Use t-distribution for confidence intervals
  mutate( ic=se * qt((1-0.01)/2 + .5, n-1))%>%
  ggplot(aes(x=subexp, y=mean)) +
  geom_bar(stat= "identity", colour="#3E3E3E", fill="#8FAADC") +
  ggtitle("Fillers")+
  ylab("Percentage of correct responses to fillers") +
  xlab("Subexperiment")+
  theme(legend.position='none')+
  geom_errorbar( aes(x=subexp, ymin=mean-se, ymax=mean+se), width=0.4, colour="#3E3E3E", alpha=0.9, linewidth=0.8)+
  coord_cartesian(ylim = c(0, 1))
```

## Exclude participants based on accuracy

```{r}
# Calculate mean accuracy per participant on fillers
agg_acc_fil <- aggregate(accuracy_num ~ id, data = fillers, mean)
names(agg_acc_fil) <- c("id", "mean")

# Define good seeds (> 75% accuracy)
good_pps <- agg_acc_fil[agg_acc_fil$mean > .75,]$id
other_pps <- agg_acc_fil[agg_acc_fil$mean <= .75,]$id

data_exp_pp <- subset(data_exp, id %in% good_pps)

cat("Excluded participants:", length(unique(data$id)) - length(unique(data_exp_pp$id)), "\n")
```

## Exclude participants based on accuracy

```{r}
# Calculate mean accuracy per participant on fillers
agg_acc_fil <- aggregate(accuracy_num ~ id, data = fillers, mean)
names(agg_acc_fil) <- c("id", "mean_acc")

id_to_group <- unique(data[, c("id", "subexp")])
agg_acc_fil <- merge(agg_acc_fil, id_to_group, by = "id")

threshold <- 0.75

good_pps <- agg_acc_fil[agg_acc_fil$mean_acc > threshold, ]$id
bad_pps <- agg_acc_fil[agg_acc_fil$mean_acc <= threshold, ]$id

exclusion_summary <- agg_acc_fil %>%
  group_by(subexp) %>%
  summarise(
    Total_Seeds = n(),
    Kept = sum(mean_acc > threshold),
    Excluded = sum(mean_acc <= threshold),
    Exclusion_Rate = paste0(round(sum(mean_acc <= threshold) / n() * 100, 1), "%")
  )

total_row <- data.frame(
  subexp = "TOTAL",
  Total_Seeds = sum(exclusion_summary$Total_Seeds),
  Kept = sum(exclusion_summary$Kept),
  Excluded = sum(exclusion_summary$Excluded),
  Exclusion_Rate = paste0(round(sum(exclusion_summary$Excluded) / sum(exclusion_summary$Total_Seeds) * 100, 1), "%")
)

final_summary <- rbind(exclusion_summary, total_row)

print(knitr::kable(final_summary, caption = "Number of Excluded Seeds by Sub-experiment (Filler Accuracy <= 75%)"))

data_exp_pp <- subset(data_exp, id %in% good_pps)

cat("\nTotal excluded participants:", length(bad_pps), "\n")
```


## Split data of included pps into primes and targets

```{r}
data_exp_pp_prime <- subset(data_exp_pp, is_prime)
data_exp_pp_target <- subset(data_exp_pp, !is_prime)

data_exp_pp_target$condition <- NA
data_exp_pp_target$condition[data_exp_pp_target$Prev_Prime_Cat == 'Critical'] <- 1
data_exp_pp_target$condition[data_exp_pp_target$Prev_Prime_Cat == 'Control'] <- 0

data_exp_pp_target$condition[data_exp_pp_target$Group == 'BAS'] <- NA

data_exp_pp_target$condition_factor <- as.factor(ifelse(data_exp_pp_target$condition==1, "critical", "control"))
```

## Analysis of Accuracy in Primes

```{r}
prime_accuracy_summary <- data_exp_pp_prime %>%
  group_by(subexp) %>%
  summarise(
    Critical_Prime_Count = sum(Prime_Category == 'Critical', na.rm = TRUE),
    Control_Prime_Count = sum(Prime_Category == 'Control', na.rm = TRUE),
    
    Critical_Prime_Acc = mean(accuracy_num[Prime_Category == 'Critical'], na.rm = TRUE),
    Control_Prime_Acc = mean(accuracy_num[Prime_Category == 'Control'], na.rm = TRUE),
    Total_Prime_Acc = mean(accuracy_num, na.rm = TRUE),
    
    n_primes = n()
  )

print(prime_accuracy_summary)
data_exp_pp_prime %>%
  filter(!is.na(Prime_Category) & Prime_Category != "") %>%
  group_by(subexp, Prime_Category) %>%
  summarise(mean_acc = mean(accuracy_num), 
            se = sd(accuracy_num)/sqrt(n()), .groups = 'drop') %>%
  ggplot(aes(x = subexp, y = mean_acc, fill = Prime_Category)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean_acc - se, ymax = mean_acc + se), 
                width = 0.2, position = position_dodge(0.9)) +
  labs(title = "Accuracy in Prime Trials by Category", y = "Accuracy", x = "Group") +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw()
```

## Plot and tabulate accuracy (Target Choice)

```{r}
# Plot Target Accuracy (Open Picture Choices / Is_Correct) by Group
data_exp_pp_target %>%
   group_by(subexp) %>%
  dplyr::summarise( 
    n=n(),
    mean=mean(accuracy_num),
    sd=sd(accuracy_num)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  mutate( ic=se * qt((1-0.01)/2 + .5, n-1))%>%
  ggplot(aes(x=subexp, y=mean)) + 
  geom_bar(stat= "identity", colour="#3E3E3E", fill="#8FAADC") +
  ggtitle("TARGET")+
  ylab("Percentage of correct responses (Target)") + 
  xlab("Subexperiment")+
  theme(legend.position='none')+
  geom_errorbar( aes(x=subexp, ymin=mean-se, ymax=mean+se), width=0.4, colour="#3E3E3E", alpha=0.9, size=0.8)+
  coord_cartesian(ylim = c(0, 1))

data_exp_pp_target %>%
   group_by(subexp) %>%
  dplyr::summarise( 
    n=n(),
    mean=mean(accuracy_num),
    sd=sd(accuracy_num)
  )
```

## Filter Targets based on Prime Accuracy

```{r}
prime_performance <- data_exp_pp_prime %>%
  dplyr::select(id, item, prime_accuracy = accuracy_num)

data_exp_pp_target <- data_exp_pp_target %>%
  left_join(prime_performance, by = c("id", "item"))

cat("Total Targets (Before filtering):", nrow(data_exp_pp_target), "\n")

data_targets_valid_prime <- data_exp_pp_target %>%
  filter(prime_accuracy == 1)

cat("Total Targets (After filtering by Prime accuracy):", nrow(data_targets_valid_prime), "\n")
cat("Excluded Trials:", nrow(data_exp_pp_target) - nrow(data_targets_valid_prime), "\n")

data_exp_pp_target <- data_targets_valid_prime
```



## GLMER for BAS (Baseline)

```{r}
data_bas <- filter(data_exp_pp_target, subexp == 'BAS')

model0_bas <- glmer(accuracy_num ~ trial_index + (1+trial_index|id) + (1|item),
               data=data_bas,
               family = binomial, control = glmerControl(optimizer = 'bobyqa'))

summary(model0_bas)

plot_model(model0_bas, type="pred", terms="trial_index")
```

## GLMER and model plots for DIST (Corresponding to 'OR' in original)

```{r}
data_dist <- filter(data_exp_pp_target, subexp == 'DIST')
data_dist <- filter(data_dist, Prev_Prime_Cat %in% c('Control', 'Critical'))
data_dist$condition <- as.factor(data_dist$condition) # 1=Critical, 0=Control

model0_dist <- glmer(accuracy_num ~ trial_index * condition + (1 + trial_index | id) + (1 | item),
               data=data_dist,
               family = binomial, control = glmerControl(optimizer = 'bobyqa'))

summary(model0_dist)

ip_dist <- interact_plot(model0_dist, pred=trial_index, modx = condition,
              plot.points = TRUE, interval = TRUE, 
              legend.main = "The Prime", 
              modx.labels = c("Cont.", "Crit.")) +
  xlab("Trial Index") + ylab("The Rate of Choosing a Zero-model") +
  ggtitle("DIS") +
  theme(legend.position="bottom")

ip_dist
ggsave("dis_gemma3_1b.png")
```


## GLMER comparison: DIST vs BAS

```{r}
data_comp_dist <- filter(data_exp_pp_target, subexp %in% c('DIST', 'BAS'))

data_comp_dist$cond_comp <- ifelse(data_comp_dist$subexp == 'BAS', 'Baseline', 
                                   ifelse(data_comp_dist$condition == 1, 'Critical', 'Control'))
data_comp_dist$cond_comp <- factor(data_comp_dist$cond_comp, levels = c('Baseline', 'Control', 'Critical'))

model_comp_dist <- glmer(accuracy_num ~ trial_index * cond_comp + (1 +trial_index| id) + (1 | item),
               data=data_comp_dist,
               family = binomial, control = glmerControl(optimizer = 'bobyqa'))

summary(model_comp_dist)
plot_model(model_comp_dist, type="pred", terms=c("trial_index", "cond_comp"))
```

## GLMER and model plots for UPP (Corresponding to 'SOME' in original)

```{r}
data_upp <- filter(data_exp_pp_target, subexp == 'UPP')
data_upp <- filter(data_upp, Prev_Prime_Cat %in% c('Control', 'Critical'))
data_upp$condition <- as.factor(data_upp$condition)

model0_upp <- glmer(accuracy_num ~ trial_index * condition + (1 + trial_index | id) + (1 | item),
               data=data_upp,
               family = binomial, control = glmerControl(optimizer = 'bobyqa'))

summary(model0_upp)

ip_upp <- interact_plot(model0_upp, pred=trial_index, modx = condition,
              plot.points = TRUE, interval = TRUE, 
              legend.main = "The Prime", 
              modx.labels = c("Cont.", "Crit.")) +
  xlab("Trial Index") + ylab("The Rate of Choosing a Zero-model") +
  ggtitle("UPP") +
  theme(legend.position="bottom")

ip_upp
ggsave("upp_gemma3_1b.png")
```

## GLMER comparison: UPP vs BAS

```{r}
data_comp_upp <- filter(data_exp_pp_target, subexp %in% c('UPP', 'BAS'))

data_comp_upp$cond_comp <- ifelse(data_comp_upp$subexp == 'BAS', 'Baseline', 
                                   ifelse(data_comp_upp$Prev_Prime_Cat == 'Critical', 'Critical', 'Control'))

data_comp_upp$cond_comp <- factor(data_comp_upp$cond_comp, levels = c('Baseline', 'Control', 'Critical'))

model_comp_upp <- glmer(accuracy_num ~ trial_index * cond_comp + (1 + trial_index | id) + (1 | item),
                        data = data_comp_upp,
                        family = binomial, control = glmerControl(optimizer = 'bobyqa'))

summary(model_comp_upp)

plot_model(model_comp_upp, type="pred", terms=c("trial_index", "cond_comp"),
           title = "Comparison: UPP vs BAS")
```

## GLMER and model plots for ESQ (Corresponding to 'SUPER' in original)

```{r}
data_esq <- filter(data_exp_pp_target, subexp == 'ESQ')
data_esq <- filter(data_esq, Prev_Prime_Cat %in% c('Control', 'Critical'))
data_esq$condition <- as.factor(data_esq$condition)

model0_esq <- glmer(accuracy_num ~ trial_index * condition + (1 + trial_index | id) + (1 | item),
               data=data_esq,
               family = binomial, control = glmerControl(optimizer = 'bobyqa'))

summary(model0_esq)

ip_esq <- interact_plot(model0_esq, pred=trial_index, modx = condition,
              plot.points = TRUE, interval = TRUE, 
              legend.main = "The Prime", 
              modx.labels = c("Cont.", "Crit.")) +
  xlab("Trial Index") + ylab("The Rate of Choosing a Zero-model") +
  ggtitle("ESQ") +
  theme(legend.position="bottom")

ip_esq
ggsave("esq_gemma3_1b.png")
```

```{r}
data_comp_esq <- filter(data_exp_pp_target, subexp %in% c('ESQ', 'BAS'))

data_comp_esq$cond_comp <- ifelse(data_comp_esq$subexp == 'BAS', 'Baseline', 
                                   ifelse(data_comp_esq$Prev_Prime_Cat == 'Critical', 'Critical', 'Control'))

data_comp_esq$cond_comp <- factor(data_comp_esq$cond_comp, levels = c('Baseline', 'Control', 'Critical'))

model_comp_esq <- glmer(accuracy_num ~ trial_index * cond_comp + (1 + trial_index | id) + (1 | item),
                        data = data_comp_esq,
                        family = binomial, control = glmerControl(optimizer = 'bobyqa'))

summary(model_comp_esq)

plot_model(model_comp_esq, type="pred", terms=c("trial_index", "cond_comp"),
           title = "Comparison: ESQ vs BAS")
```

## Combining Plots

```{r}
grid.arrange(ip_dist, ip_upp, ip_esq, ncol=3)
```

## Plot priming effect across sub-experiments

```{r}
summary_plot <- data_exp_pp_target %>%
  filter(!is.na(condition_factor)) %>%
  group_by(subexp, condition_factor) %>%
  summarise(mean_acc = mean(accuracy_num), 
            se = sd(accuracy_num)/sqrt(n()),
            .groups = 'drop')

ggplot(summary_plot, aes(x = condition_factor, y = mean_acc, group = subexp, color = subexp)) +
  geom_point(size = 4) +
  geom_line(size = 1.2) +
  geom_errorbar(aes(ymin = mean_acc - se, ymax = mean_acc + se), width = 0.1) +
  labs(title = "Priming Effect Across Sub-experiments",
       x = "Condition", y = "Acceptance Rate / Accuracy",
       color = "Group") +
  theme_bw()
```

```{r}
target_acc_summary <- data_exp_pp_target %>%
  filter(condition_factor %in% c("critical", "control")) %>%
  group_by(subexp, condition_factor) %>%
  summarise(
    n = n(),
    mean_acc = mean(accuracy_num, na.rm = TRUE),
    sd = sd(accuracy_num, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    se = sd / sqrt(n) 
  )

print(knitr::kable(target_acc_summary, digits = 3, caption = "Target Accuracy by Prime Condition"))

ggplot(target_acc_summary, aes(x = subexp, y = mean_acc, fill = condition_factor)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7, color = "black", alpha = 0.8) +
  
  geom_errorbar(aes(ymin = mean_acc - se, ymax = mean_acc + se),
                position = position_dodge(width = 0.8), width = 0.25, size = 0.8) +
  
  scale_fill_manual(values = c("control" = "#999999", "critical" = "#56B4E9"),
                    labels = c("Control Prime", "Critical Prime")) +
  
  labs(
    title = "Target Accuracy: Critical vs. Control Prime",
    subtitle = "Only targets following correct primes included",
    x = "Sub-experiment",
    y = "Mean Target Accuracy",
    fill = "Condition"
  ) +
  
  coord_cartesian(ylim = c(0, 1)) +
  
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 13, face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(face = "bold")
  )
```

```{r}
data_firth <- data_exp_pp_target %>%
  filter(condition_factor %in% c("control", "critical")) %>%
  mutate(
    condition_factor = factor(condition_factor, levels = c("control", "critical")),
    subexp = factor(subexp)
  )

cat("=== Overall Firth Logistic Regression ===\n")
model_firth_all <- logistf(accuracy_num ~ condition_factor * subexp, data = data_firth)
print(summary(model_firth_all))

cat("\n=== Firth Regression: DIST ===\n")
data_dist_firth <- filter(data_firth, subexp == 'DIST')
if(nrow(data_dist_firth) > 0) {
  model_firth_dist <- logistf(accuracy_num ~ condition_factor, data = data_dist_firth)
  print(summary(model_firth_dist))
}

cat("\n=== Firth Regression: UPP ===\n")
data_upp_firth <- filter(data_firth, subexp == 'UPP')
if(nrow(data_upp_firth) > 0) {
  model_firth_upp <- logistf(accuracy_num ~ condition_factor, data = data_upp_firth)
  print(summary(model_firth_upp))
}

cat("\n=== Firth Regression: ESQ ===\n")
data_esq_firth <- filter(data_firth, subexp == 'ESQ')
if(nrow(data_esq_firth) > 0) {
  model_firth_esq <- logistf(accuracy_num ~ condition_factor, data = data_esq_firth)
  print(summary(model_firth_esq))
}

```
